---
title: "Covid-19 analysis"
author: "Caleb Moses"
date: "`r format(Sys.Date(), '%d %B %Y')`"
knit: "bookdown::render_book"
output:
    bookdown::gitbook:
        self_contained: false
        pandoc_args: ["--verbose", "--log=pandoc.log"]
documentclass: book
link-citations: yes
github-repo: mathematiguy/covid-19
description: "This report analyses publicly available covid 19 data"
---

# Introduction

```{r}

library(yaml)
library(here)
library(tidyverse)
library(data.table)

source(here('R/get_odata_fun.R'), local=TRUE)
source(here('R/get_odata_catalogue_fun.R'), local=TRUE)
options(tibble.width = Inf, mc.cores = parallel::detectCores()-1)

credentials = read_yaml(here('credentials.yaml'))

# Set default ggplot theme
theme_set(
    theme_minimal() +
    theme(plot.title = element_text(hjust = 0.5)) +
    theme(text = element_text(size = 20))
)

# Set notebook options
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, progress=TRUE, verbose=TRUE, cache.extra = knitr::rand_seed)

```

I want to build an SIR model for the covid 19 response but I need the right data in order to do it.

To get started I need the following datasets:

- NZ Population by region + ethnicity + age
- Daily case counts by region + ethnicity (at least) + age (preferably) [check]
- Hospitalisations by region + ethnicity (at least) + age (preferably)
- Deaths by region + ethnicity (at least) + age (preferably)

Lets start this by selecting a simple subset of the data that will be easier to analyse. 
Then we can chuck it into julia and see what we can do there.

```{r}

```
