---
title: "Covid-19 analysis"
author: "Caleb Moses"
date: "`r format(Sys.Date(), '%d %B %Y')`"
knit: "bookdown::render_book"
output:
    bookdown::gitbook:
        self_contained: false
        pandoc_args: ["--verbose", "--log=pandoc.log"]
documentclass: book
link-citations: yes
github-repo: mathematiguy/covid-19
description: "This report analyses publicly available covid 19 data"
---

# Introduction

```{r}

library(yaml)
library(here)
library(rstan)
library(tidyverse)
library(data.table)

source(here('R/get_odata_fun.R'), local=TRUE)
source(here('R/get_odata_catalogue_fun.R'), local=TRUE)
options(tibble.width = Inf, mc.cores = parallel::detectCores()-1)

credentials = read_yaml(here('credentials.yaml'))

# Set default ggplot theme
theme_set(
    theme_minimal() +
    theme(plot.title = element_text(hjust = 0.5)) +
    theme(text = element_text(size = 20))
)

# Set notebook options
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, progress=TRUE, verbose=TRUE, cache.extra = knitr::rand_seed)

```

I want to build an SIR model for the covid 19 response but I need the right data in order to do it.

To get started I need the following datasets:

- NZ Population by region + ethnicity + age
- Daily case counts by region + ethnicity (at least) + age (preferably) [check]
- Hospitalisations by region + ethnicity (at least) + age (preferably)
- Deaths by region + ethnicity (at least) + age (preferably)

Lets start this by selecting a simple subset of the data that will be easier to analyse. 
Then we can chuck it into julia and see what we can do there.

```{r}

auckland_cases <- read_csv(here('data/auckland_cases.csv'))

# Select a subsample of data where it looks like an SIR model would apply
sir_data <- auckland_cases %>%
    filter(Date < as.Date('2021-09-21')) %>%
    mutate(S = 5000000 - Total,
           I = Active,
           R = Recovered + Deceased) %>%
    select(Date, S, I, R)

sir_data %>%
    gather(S:R, key = 'Compartment', value = 'Count') %>%
    mutate(Compartment = factor(Compartment, levels = c('S', 'I', 'R'))) %>%
    ggplot(aes(x = Date, y = Count, colour = Compartment)) +
    geom_line() +
    facet_wrap(~Compartment, ncol = 1, scales = 'free_y')

```

```{r}

# time series of cases
cases <- sir_data$I

# total count
N <- 5000000;

# times
n_days <- as.numeric(sir_data$Date[nrow(sir_data)] - sir_data$Date[1])
t <- seq(0, n_days, by = 1)
t0 = 0
t <- t[-1]

#initial conditions
i0 <- 1
s0 <- N - i0
r0 <- sir_data$R[1]
y0 = c(S = s0, I = i0, R = r0)

# data for Stan
data_sir <- list(n_days = n_days, y0 = y0, t0 = t0, ts = t, N = N, cases = cases)

# number of MCMC steps
niter <- 2000

```

```{r}

# Compile the model
model <- stan_model(here("stan/sir_model.stan"))

fit_sir_negbin <- sampling(
    model,
    data = data_sir,
    iter = niter,
    chains = 1,
    seed = 0)

```
