---
title: "Covid-19 analysis"
author: "Caleb Moses"
date: "`r format(Sys.Date(), '%d %B %Y')`"
knit: "bookdown::render_book"
output:
  bookdown::gitbook:
    self_contained: false
    pandoc_args: ["--verbose", "--log=pandoc.log"]
documentclass: book
link-citations: yes
github-repo: mathematiguy/covid-19
description: "This report analyses publicly available covid 19 data"
---

# Building the branching process model

On this page we focus on creating all of the pieces we need to reconstruct the model.

```{r setup}

library(tidyverse)

# Set default ggplot theme
theme_set(
    theme_minimal() +
    theme(plot.title = element_text(hjust = 0.5, size = 15)) +
    theme(text = element_text(size = 12))
)

numsims = 1000

```

## Model parameters

We can build each of the model pieces one by one. Once we have the coded up, we can hopefully write an implementation in Stan.

### $Y_l$

$Y_l$ is gamma distributed with a mean of 1 and a variance of $1/k$ for $k$ = 0.5. Hence the variance of $Y_l$ is equal to 2.

```{r Y_l}

Y_l = rgamma(100000, shape=0.5, scale=2)

tibble(Y_l=Y_l) %>%
    ggplot(aes(x=Y_l)) +
    geom_density(alpha=0.4, fill='#FF6666') +
    ggtitle(str_wrap("Distribution of individual level heterogeneity", 50))

cat("The sample mean of Y_l is:", mean(Y_l), '\n')
cat("The sample variance of Y_l is:", var(Y_l), '\n')

```

### $V_l(t)$

```{r set-vaccine-transmission-effects}
e_T1 = 0
e_T2 = 0.5
```

$V_l(t)$ represents the effect of vaccination on the transmission rate of individual $l$. It depends on 2 values, $e_{T1}$ = `r e_T1` and $e_{T2}$ = `r e_T2`.

```{r V_l}

V_l <- tibble(doses = c(0, 1, 2), eff_transmission = c(1, 1 - e_T1, 1 - e_T2))

V_l %>%
    mutate(doses = factor(doses)) %>%
    ggplot(aes(x = doses, y = eff_transmission)) +
    geom_bar(stat = 'identity', width = 0.01) +
    geom_point(size=5) +
    ylim(0,1) +
    ggtitle(str_wrap('Number of vaccine doses vs vaccine effectiveness against transmission', 50))

```

### $F_l(t)$

$F_l(t)$ represents the effect of quarantine or isolation on the transmission rate of individual $l$ at time $t$.

```{r F_l}

c_quar = 0.5
c_isol = 0

F_l <- tibble(quarantine = c('Not in quarantine/isolation', 'In quarantine', 'In isolation'), eff_transmission = c(1, c_quar, c_isol))

F_l %>%
    mutate(quarantine = factor(quarantine)) %>%
    ggplot(aes(x = quarantine, y = eff_transmission)) +
    geom_bar(stat = 'identity', width = 0.01) +
    geom_point(size=5) +
    ylim(0,1) +
    ggtitle(str_wrap('Effectiveness against transmission vs Quarantine/Isolation status', 50))

```

### $w(\tau)$

$w(\tau)$ is the probability density function of the assumed generation time distribution and $t_{inf,l}$ is the time individual $l$ was infected. In the absence of interventions, we assume generation times are drawn from a Weibull distribution with mean 5.0 days and s.d. 1.9 days.

```{r w_tau}

# Consider replacing grid search with: https://cran.r-project.org/web/packages/rootSolve/vignettes/rootSolve.pdf
# Use grid search to find the lambda + k params give us the mean + variance that we want
## w_scan = tibble(lambda = seq(1, 10, 0.001)) %>%
##     mutate(k = map(lambda, function(x) seq(1, 10, 0.01))) %>%
##     unnest(k) %>%
##     mutate(mean = map2_dbl(lambda, k, function(l, k) l * gamma(1 + 1/k))) %>%
##     mutate(var = map2_dbl(lambda, k, function(l, k) l^2 * (gamma(1+2/k) - gamma(1+1/k)^2))) %>%
##     mutate(mean_error = abs(mean - 5),
##            var_error = abs(var - 1.9 ^ 2),
##            total_error = (mean_error + var_error) / 2) %>%
##     arrange(total_error) %>%
##     head()

w_tau = rweibull(10000, shape = 2.85, scale = 5.604)

tibble(w_tau=w_tau) %>%
    ggplot(aes(x=w_tau)) +
    geom_density(alpha=0.4, fill='#FF6666') +
    ggtitle("Generation time distribution")

cat("The sample mean of w(tau) is:", mean(w_tau), '\n')
cat("The sample variance of w(tau) is:", var(w_tau), '\n')
cat("The sample standard deviation of w(tau) is:", sd(w_tau), '\n')

```

### $NGM^{clin}$

$NGM^{clin}$ is the next generation matrix for clinical individuals and $a_l$ is the age group of individual $l$. We use a contact matrix sourced from Prem et al.

```{r build-ngm-matrix}

# Age group labels
age_groups = c("0-4", "5-9", "10-14", "15-19", "20-24", "25-29", "30-34", "35-39", "40-44",
               "45-49", "50-54", "55-59", "60-64", "65-69", "70-74", "75+")

# Get Age specific model parameters
age_params <- read_csv(here('data/age_specific_parameters.csv')) %>%
    gather(`0-4`:`75+`, key = 'Age_Group', value = 'Value') %>%
    spread(key = Age, value = Value) %>%
    mutate(Age_Group = factor(Age_Group, levels = age_groups)) %>%
    arrange(Age_Group)

# Grab variables necessary for NGM matrix calculation
p_clin = age_params$`Pr(clinical)`
u = age_params$Susceptibility
N = age_params$`% of popn`

# Load the contact matrix
M <- here('data/contact_matrix.csv') %>%
    read_csv() %>%
    as.matrix()

# Build the NGM matrix
R_0 = 6.0
tau = 0.5
NGM = matrix(0, dim(M)[2], dim(M)[1])
for (i in 1:16) {
    for (j in 1:16) {
        NGM[i,j] = ( p_clin[j] + tau * (1 - p_clin[j]) ) * u[i] * (M[i,j] + N[j] / N[i] * M[j,i]) / 2
    }
}
NGM <- NGM * (R_0 / max(eigen(NGM)$values))
colnames(NGM) <- age_groups

# Display the matrix as a heatmap
NGM  %>%
    as_tibble() %>%
    mutate(Age_Group_1 = age_groups) %>%
    gather(`0-4`:`75+`, key = "Age_Group_2", value = "Contact Rate") %>%
    mutate(Age_Group_1 = factor(Age_Group_1, levels = age_groups),
           Age_Group_2 = factor(Age_Group_2, levels = age_groups)) %>%
    ggplot(aes(x = Age_Group_2, y = Age_Group_1, fill = `Contact Rate`)) +
    geom_tile() +
    ggtitle("Next generation matrix by age group") +
    xlab("") +
    ylab("") +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

```

### $s_j(t)$

$s_j(t)$ is the fraction of age group $j$ that is susceptible at time $t$

```{r s_j, echo=TRUE}

s <- function(j, e_I1, e_I2, c_1, c_2, N, N_tot) {
    # e_I1, e_I2 is the effectiveness of the vaccine against infection after 1 and 2 doses respectively
    # c_1, c_2 is the fraction of age group j that has received d effective doses of the vaccine at time t
    # N is the cumulative number of infections in age group j
    # N_tot is the total size of age group j
    return(1 - e_I1*c_1 - e_I2*c_2 - N / N_tot)
}

```

## Prior predictive check

Now that we have each piece of the model built, we can perform a 'prior predictive check'. That is, we can sample from the prior distribution for the model to see what the results look like.


```{r}

```
