---
title: Analysis of COVID-19 cases in New Zealand
author: Caleb Moses
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
    prettydoc::html_pretty:
      theme: tactile
highlight: github
---

```{r import-libraries, message = FALSE, echo = FALSE}
# Import libraries
library(tidyverse)
library(nCov2019)
library(wbstats)
library(broom)
library(furrr)
library(rstan)
library(brms)
library(here)

# Use furr multiprocessing
plan(multiprocess)

# Set default ggplot theme
theme_set(theme_minimal() + theme(plot.title = element_text(hjust = 0.5)))

# Set up stan options
options(mc.cores = parallel::detectCores())
rstan_options(auto_write = TRUE)

knitr::opts_chunk$set(echo = FALSE, fig.align = "center", warning = FALSE, message = FALSE, fig.width = 12)
```

## Get the data

We fetch the latest covid-19 data using the `nCov2019` R package.

```{r get-data}

# Get global COVID-19 data
covid_data <- load_nCov2019(lang = 'en')['global'] %>%
    filter(country == "New Zealand") %>%
    group_by(time, country) %>%
    summarise(cum_confirm = sum(cum_confirm),
              cum_heal = sum(cum_heal),
              cum_dead = sum(cum_dead)) %>%
    ungroup()

```

## Visualise the data

Here we plot the number of confirmed cases over time.

```{r cases-over-time}

# Plot the number of cases over time
covid_data %>%
    ggplot(aes(x = time, y = cum_confirm)) +
    geom_point() +
    geom_line() +
    ylab("Cumulative confirmed cases") +
    ggtitle("Cases of COVID-19 in New Zealand over time")

```

## Model the data

We need to make some assumptions in order to build a simulation model of the covid-19
cases. We are interested in forecasting the future cases, calculating the basic
reproduction number and combining a variety of related data sources to this problem.

### The exponential model

We can work towards this by building a sequence of increasingly complex models. The
most simple of these being a simple exponential model.

The model stan code is shown below. It implements a simple discrete model with two parameters,
a level and a slope `alpha` are used to define an exponential model. The output of the model 
`y` is a continuous parameter.

#### The continuous prior model

```{bash exp-prior-stan}
cat exponential_model.stan
```

```{r exponential-prior, results = 'hide'}

# Prior simulation
exponential_prior <- stan(
    file = here("analysis/exponential_prior.stan"),
    data = list(
        N = nrow(covid_data),
        x = 1:nrow(covid_data),
        y = covid_data$cum_confirm),
    iter = 2000,
    chains = 4,
    algorithm = "Fixed_param"
)

y_prior = extract(exponential_prior) %>%
    .$y_prior %>%
    as_tibble() %>%
    add_column(sim = 1:nrow(.), .before = "V1") %>%
    gather(paste0("V", 1:nrow(covid_data)), key = 'time', value = 'cum_confirm') %>%
    mutate(time = as.integer(str_remove_all(time, "V")) + min(covid_data$time) - 1)

prior_plot <- y_prior %>%
    ggplot(aes(x = time, y = cum_confirm, group = sim)) +
    geom_line(alpha = 0.1, aes(colour = 1)) +
    guides(colour = FALSE) +
    ylab("Simulated cumulative cases") +
    ggtitle("Prior simulated cumulative cases of COVID-19 in NZ over time")

prior_plot

```

#### The continuous posterior model

```{bash exp-posterior-stan, echo = FALSE}
cat exponential_model.stan
```

```{r exponential-model, results = 'hide'}

# Prior simulation
exponential_post <- stan(
    file = here("analysis/exponential_model.stan"),
    data = list(
        N = nrow(covid_data),
        y = covid_data$cum_confirm),
    iter = 2000,
    chains = 4
)

exponential_post

y_post = extract(exponential_post) %>%
    .$y_pred %>%
    as_tibble() %>%
    add_column(sim = 1:nrow(.), .before = "V1") %>%
    gather(paste0("V", 1:nrow(covid_data)), key = 'time', value = 'cum_confirm') %>%
    mutate(time = as.integer(str_remove_all(time, "V")) + min(covid_data$time) - 1)

y_post %>%
    filter(time == max(time)) %>%
    .$cum_confirm %>%
    quantile(probs = c(0.01, 0.05, 0.25, 0.5, 0.75, 0.95, 0.99))

```

The plot below shows the posterior distribution of cumulative expected cases, conditioned on the
observed data.

```{r exp-prior-data-posterior-plot}

y_prior %>%
    rename(y_prior = cum_confirm) %>%
    left_join(rename(y_post, y_post = cum_confirm), by = c('sim', 'time')) %>%
    gather(y_prior:y_post, key = 'variable', value = 'value') %>%
    mutate(variable = case_when(
               variable == "y_prior" ~ "Prior",
               variable == "y_data" ~ "Data",
               variable == "y_post" ~ "Posterior")) %>%
    mutate(variable = factor(variable, levels=c('Prior', 'Data', 'Posterior'))) %>%
    ggplot(aes(x = time, y = value, group = sim, colour = variable)) +
    geom_line(alpha = 0.01) +
    geom_point(data = covid_data %>%
                   rename(value = cum_confirm) %>%
                   mutate(sim = 1, variable = 'y_data') %>%
                   select(sim, time, variable, value) %>%
                   mutate(variable = case_when(
                              variable == "y_prior" ~ "Prior",
                              variable == "y_data" ~ "Data",
                              variable == "y_post" ~ "Posterior")) %>%
                   mutate(variable = factor(variable, levels=c('Prior', 'Data', 'Posterior')))) +
    facet_wrap(~variable) +
    guides(colour = FALSE) +
    ylim(0, max(covid_data$cum_confirm) * 2) +
    ylab("Expected number of cases") +
    ggtitle("Prior, observed and posterior count of covid-19 cases in NZ over time")

y_prior %>%
    rename(y_prior = cum_confirm) %>%
    left_join(rename(y_post, y_post = cum_confirm), by = c('sim', 'time')) %>%
    gather(y_prior:y_post, key = 'variable', value = 'value') %>%
    mutate(variable = case_when(
               variable == "y_prior" ~ "Prior",
               variable == "y_data" ~ "Data",
               variable == "y_post" ~ "Posterior")) %>%
    mutate(variable = factor(variable, levels=c('Prior', 'Data', 'Posterior'))) %>%
    ggplot(aes(x = time, y = value, group = sim, colour = variable)) +
    geom_line(alpha = 0.01) +
    geom_point(data = covid_data %>%
                   rename(value = cum_confirm) %>%
                   mutate(sim = 1, variable = 'y_data') %>%
                   select(sim, time, variable, value) %>%
                   mutate(variable = case_when(
                              variable == "y_prior" ~ "Prior",
                              variable == "y_data" ~ "Data",
                              variable == "y_post" ~ "Posterior")) %>%
                   mutate(variable = factor(variable, levels=c('Prior', 'Data', 'Posterior')))) +
    facet_wrap(~variable) +
    guides(colour = FALSE) +
    scale_y_log10() +
    ylab("Expected number of cases")

```

This plot shows the prior and posterior distributions of the model parameters (`alpha` and `beta`). 

```{r prior-vs-post-params}

param_data <- bind_rows(
  tibble(
      prior_alpha = extract(exponential_prior)$alpha,
      post_alpha = extract(exponential_post)$alpha) %>%
      mutate(sims = 1:nrow(.)) %>%
      gather(prior_alpha:post_alpha, key = 'variable', value = 'value'),
  tibble(
      prior_beta = extract(exponential_prior)$y_prior[,1],
      post_beta = extract(exponential_post)$y[,1]) %>%
      mutate(sims = 1:nrow(.)) %>%
  gather(prior_beta:post_beta, key = 'variable', value = 'value')) %>%
    separate(variable, into = c('model', 'variable')) %>%
    mutate(model = factor(
               if_else(model == 'prior', 'Prior', 'Posterior'),
               levels = c('Prior', 'Posterior')),
           variable = factor(if_else(variable == 'alpha', 'Alpha', 'Beta'),
                             levels = c('Alpha', 'Beta')))

param_data %>%
    ggplot(aes(x = value, fill = variable)) +
    geom_histogram() +
    facet_grid(variable~model) +
    guides(fill = FALSE)

param_data %>%
    spread(key = variable, value = value) %>%
    ggplot(aes(x = Alpha, y = Beta, colour = model)) +
    facet_wrap(~model) +
    geom_point(alpha = 0.2) +
    guides(colour = FALSE)

```

### The logistic model

Now we can prepare a logistic model. This requires us to include the total population into the model.
As the count of cases approaches the total population, the growth will begin to slow.

```{r}

# Get global population data
pop_data <- wb(indicator = "SP.POP.TOTL", startdate = 1900, enddate = 2019) %>%
    filter(country == "New Zealand") %>%
    mutate_at(c("date", "value"), as.integer)

pop_model <- lm(value ~ 1 + date, data = pop_data)
pop_estimate <- as.integer(predict(pop_model, newdata = data.frame(date = 2020)))
```

#### Prior simulation

We begin again with a prior simulation. This time we may care about the short and the long term
behaviour just to make sure it is tuned properly.

```{r logistic-prior}

# Prior simulation
logistic_prior <- stan(
    file = here("analysis/logistic_prior.stan"),
    data = list(
        N = nrow(covid_data),
        K = 5000),
    iter = 2000,
    chains = 4,
    algorithm = "Fixed_param"
)

y_prior = extract(logistic_prior) %>%
    .$y_prior %>%
    as_tibble() %>%
    add_column(sim = 1:nrow(.), .before = "V1") %>%
    gather(paste0("V", 1:nrow(covid_data)), key = 'time', value = 'cum_confirm') %>%
    mutate(time = as.integer(str_remove_all(time, "V")) + min(covid_data$time) - 1)

prior_plot <- y_prior %>%
    ggplot(aes(x = time, y = cum_confirm, group = sim)) +
    geom_line(alpha = 0.1, aes(colour = 1)) +
    guides(colour = FALSE) +
    ylab("Simulated cumulative cases") +
    ggtitle("Prior simulated cumulative cases of COVID-19 in NZ over time")

prior_plot

```

```{r logistic-post}

# Prior simulation
logistic_post <- stan(
    file = here("analysis/logistic_model.stan"),
    data = list(
        N = nrow(covid_data),
        y = covid_data$cum_confirm,
        K = 5000),
    iter = 2000,
    chains = 4
)

y_post = extract(logistic_post) %>%
    .$y_pred %>%
    as_tibble() %>%
    add_column(sim = 1:nrow(.), .before = "V1") %>%
    gather(paste0("V", 1:nrow(covid_data)), key = 'time', value = 'cum_confirm') %>%
    mutate(time = as.integer(str_remove_all(time, "V")) + min(covid_data$time) - 1)

y_prior %>%
    rename(y_prior = cum_confirm) %>%
    left_join(rename(y_post, y_post = cum_confirm), by = c('sim', 'time')) %>%
    gather(y_prior:y_post, key = 'variable', value = 'value') %>%
    mutate(variable = case_when(
               variable == "y_prior" ~ "Prior",
               variable == "y_data" ~ "Data",
               variable == "y_post" ~ "Posterior")) %>%
    mutate(variable = factor(variable, levels=c('Prior', 'Data', 'Posterior'))) %>%
    ggplot(aes(x = time, y = value, group = sim, colour = variable)) +
    geom_line(alpha = 0.01) +
    geom_point(data = covid_data %>%
                   rename(value = cum_confirm) %>%
                   mutate(sim = 1, variable = 'y_data') %>%
                   select(sim, time, variable, value) %>%
                   mutate(variable = case_when(
                              variable == "y_prior" ~ "Prior",
                              variable == "y_data" ~ "Data",
                              variable == "y_post" ~ "Posterior")) %>%
                   mutate(variable = factor(variable, levels=c('Prior', 'Data', 'Posterior')))) +
    facet_wrap(~variable) +
    guides(colour = FALSE) +
    ylim(0, max(covid_data$cum_confirm) * 2) +
    ylab("Expected number of cases") +
    ggtitle("Prior, observed and posterior count of covid-19 cases in NZ over time")

```

```{r logistic-param-plot}

param_data <- bind_rows(
    tibble(
        prior_alpha = extract(exponential_prior)$alpha,
        post_alpha = extract(exponential_post)$alpha) %>%
    mutate(sims = 1:nrow(.)) %>%
    gather(prior_alpha:post_alpha, key = 'variable', value = 'value'),
    tibble(
        prior_beta = extract(exponential_prior)$y_prior[,1],
        post_beta = extract(exponential_post)$y[,1]) %>%
    mutate(sims = 1:nrow(.)) %>%
    gather(prior_beta:post_beta, key = 'variable', value = 'value')) %>%
    separate(variable, into = c('model', 'variable')) %>%
    mutate(model = factor(
               if_else(model == 'prior', 'Prior', 'Posterior'),
               levels = c('Prior', 'Posterior')),
           variable = factor(if_else(variable == 'alpha', 'Alpha', 'Beta'),
                             levels = c('Alpha', 'Beta')))

param_data %>%
    ggplot(aes(x = value, fill = variable)) +
    geom_histogram() +
    facet_grid(variable~model) +
    guides(fill = FALSE)

param_data %>%
    spread(key = variable, value = value) %>%
    ggplot(aes(x = Alpha, y = Beta, colour = model)) +
    facet_wrap(~model) +
    geom_point(alpha = 0.2) +
    guides(colour = FALSE)

```

```{r}

num_timesteps = nrow(covid_data) + 5

extrapolate_logistic_model <- function(alpha, beta, timesteps, K) {

    y <- rep(0, timesteps)
    y[1] = beta
    for (i in 2:timesteps) {
        y[i] = (1 + alpha) * y[i-1] * ( 1 - y[i-1] / K)
    }
    return(y)
}

extrapolate_data <- param_data %>%
    filter(model == "Posterior") %>%
    select(-model) %>%
    spread(key = variable, value = value) %>%
    mutate(y = map2(Alpha, Beta,
                    function(a, b) extrapolate_logistic_model(a, b, num_timesteps, pop_estimate))) %>%
    unnest(y) %>%
    mutate(time = rep(1:num_timesteps, 4000) + min(covid_data$time) - 1)

alpha_quantiles = quantile(extract(logistic_post)$alpha, probs = c(0.05, 0.5, 0.95))

today_cases = covid_data$cum_confirm[nrow(covid_data)]

today_cases * (alpha_quantiles - 1)

```


```{r softplus-prior}

# Prior simulation
softplus_prior <- stan(
    file = here("analysis/softplus_prior.stan"),
    data = list(
        N = nrow(covid_data),
        K = 5000),
    iter = 2000,
    chains = 4,
    algorithm = "Fixed_param"
)

y_prior = extract(softplus_prior) %>%
    .$y_prior %>%
    as_tibble() %>%
    add_column(sim = 1:nrow(.), .before = "V1") %>%
    gather(paste0("V", 1:nrow(covid_data)), key = 'time', value = 'cum_confirm') %>%
    mutate(time = as.integer(str_remove_all(time, "V")) + min(covid_data$time) - 1)

y_prior$cum_confirm %>% max()

prior_plot <- y_prior %>%
    ggplot(aes(x = time, y = cum_confirm, group = sim)) +
    geom_line(alpha = 0.1, aes(colour = 1)) +
    guides(colour = FALSE) +
    ylab("Simulated cumulative cases") +
    ggtitle("Prior simulated cumulative cases of COVID-19 in NZ over time")

prior_plot

```
