---
title: A stochastic model for COVID-19 spread based on work by Te Punaha Matatini
author: Caleb Moses
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
    prettydoc::html_pretty:
      theme: tactile
highlight: github
---

```{r import-libraries, message = FALSE, echo = FALSE}

# Import libraries
library(tidyverse)
library(nCov2019)
library(wbstats)
library(broom)
library(furrr)
library(rstan)
library(brms)
library(here)

# Use furr multiprocessing
plan(multiprocess)

# Set default ggplot theme
theme_set(theme_minimal() + theme(plot.title = element_text(hjust = 0.5)))

# Set up stan options
options(mc.cores = parallel::detectCores())
rstan_options(auto_write = TRUE)

knitr::opts_chunk$set(echo = FALSE, fig.align = "center", warning = FALSE, message = FALSE, fig.width = 12)

```

# Introduction

This notebook implements a discrete-time branching process model based on the paper "A stochastic model for
COVID-19 spread and the effects of Alert Level 4 in Aotearoa New Zealand", released on the 9th of April by
Te Pūnaha Matatini, using the Stan probabilistic programming language.

## Load the data

```{r}

calculate_new_cases <- function(x) {
    lag_x <- lag(x)
    last_x <- ifelse(is.na(lag_x), 0, lag_x)
    x - last_x
}

# Get NZ COVID-19 data
covid_data <- load_nCov2019(lang = 'en')['global'] %>%
    filter(country == "New Zealand") %>%
    group_by(time, country) %>%
    summarise_at(vars(cum_confirm:cum_dead), sum) %>%
    ungroup() %>%
    group_by(country) %>%
    mutate(new_confirm = calculate_new_cases(cum_confirm),
           new_heal = calculate_new_cases(cum_heal),
           new_dead = calculate_new_cases(cum_dead)) %>%
    ungroup()

covid_data %>%
    gather(starts_with("new"), key = 'variable', value = 'value') %>%
    ggplot(aes(x = time, y = value, colour = variable)) +
    geom_point() +
    facet_wrap(~variable, ncol = 1)

sir_data <- covid_data %>%
    mutate(S = as.integer(5000000 - cum_confirm - cum_heal - cum_dead),
           I = as.integer(cum_confirm - cum_heal - cum_dead),
           R = as.integer(cum_heal + cum_dead)) %>%
    select(time, S, I, R)

sir_data %>%
    gather(S:R, key = 'variable', value = 'value') %>%
    ggplot(aes(x = time, y = value, colour = variable)) +
    geom_point() +
    facet_wrap(~variable, scales = 'free_y') +
    scale_y_log10()

```

# The discrete time branching process model

The key model assumptions are:

## Infected individuals

Infected individuals are grouped into two categories:

1. Those who show clinical symptoms at some point during their infection
2. Those who are subclinical

## Self-Isolation measures

In the absence of self isolation measures, each infected individual causes a randomly generated number
$N_i ~ Poisson(R_i)$ of new infections.

For clinical individuals, $R_i = R_{clin}$ and for subclinical individuals, $R_i = R_{sub}$, and we assume
$R_{sub} = 0.5 R_{clin}$.

## Time to infection

The time between an individual becoming infected and infecting another individual, called the generation
time $T_G$ follows a Weibull distribution with mean and median equal to 5.0 days, and standard deviation
of 1.9 days.

The infection times of all $N_i$ secondary infections from each individual are independent, identically
distributed random variables.

## Asymptomatic period

Clinical individuals have an initial period during which they are either asymptomatic or have sufficiently
mild symptoms that they have not self-isolated. During this period their infectiousness is as shown
by the blue curve in figure 1.

At the end of this period, once they have developed more serious symptoms, they are isolated and their
infectiousness reduces to $c_{iso} = 65\%$ of the value it would have without isolation.

This represents a control policy of requiring symptomatic individuals to self-isolate.

## Subclinical individuals

Subclinical individuals do not get isolated and are not reported in case data.

## Recovery

All individuals are assumed to be no longer infectious 30 days after being infected. This is an
upper limit for computational convenience; in practice, individuals have very low infectiousness
after about 12 days because of the shape of the genration time distribution.

Individuals who have recovered from the virus are assumed to have immunity and cannot be infected
again. This means that the proportion of the population that is susceptable at time $t$ is
$1 - N(t) / N_{pop}$, where $N(t)$ is the cumulative number of infections at time $t$ and $N_{pop}$
is the total population size.

## Time between infection and isolation

The time $T_{iso}$ between infection and isolation is the sum of two random variables $T_1$ and $T_2$.
$T_1$ represents the incubation period (time from infection to onset of symptoms) and has a gamma
distribution with mean 5.5 days and shape parameter 5.7. $T_2$ represents the time from onset to
isolation and is taken from the New Zealand case data.

## Model equation

The model is simulated using a timestep of $\delta t = 1$ day. At each step, infectious individual $i$
produces a Poisson distributed number of secondary infections with mean:

$$\lambda_i = R_i \left(1 - \frac{N(t)}{N_{pop}}\right) C(t) F\left(t - T_{I,i} - T_{iso,i} \right) \int_t^{t+\delta t} W \left(\tau - T_{I,i}\right) d\tau$$

where $R_i \in \{R_{clin}, R_{sub}\}$ is the individual's mean number of secondary infections, $T_{I,i}$
is the time individual $i$ became infected, $T_{iso,i}$ is the delay from becoming infected to being
isolated, $C(t)$ is the control effectivity at time $t$ and $F(t)$ is a function describing the reduction
in infectiousness due to isolation:

\[F(t) = \begin{cases}
      1 & s < 0 \\
      c_{iso} & s > 0
   \end{cases}
\]

## Clinical infections

Clinical infections have a probability $p_H = 7.8\%$ of being hospitalised, equivalent to an overall
infection-hospitalisation ration of 5.25%. The duration of hospital stay is exponentially distributed
with mean 10 days. For simplicity, hospitalisation is assumed to occur at the same time as onset of
symptoms, i.e. time $T_1$ after infection.

## Seed cases

Te Pūnaha Matatini initialised their model with seed cases representing the arrival of actual infected
individuals from overseas. The number and timing of these seed cases was chosen to replicate real case
data.

Since we do not have this precise data(?) we have to initialise our seed cases with a model of
international arrivals.

# Prior simulation

We begin with a stan model so simulate the prior distribution of the outputs. This will help us to test
simple scenarios, to make sure the model is working properly.

## Simulating cases

```{r}

# Prior simulation
sir_model <- stan(
    file = here("analysis/tmp_stochastic_prior.stan"),
    data = list(
        N = 5000000,
        T = nrow(sir_data)
    ),
    iter = 2000,
    chains = 4,
    algorithm = 'Fixed_param'
)

sir_model

```

## Visualising the results

```{r}

clean_up_posterior <- function(post, var_name) {
    post %>%
        as_tibble() %>%
        add_column(sims = 1:nrow(.), .before = 'V1') %>%
        gather(starts_with("V"), key = 'time', value = !!var_name) %>%
        mutate(time = as.integer(str_remove(time, "V")))
}

var_names = c('S', 'I', 'R')

posterior <- map2(extract(sir_model)[var_names], var_names, clean_up_posterior) %>%
    reduce(left_join) %>%
    mutate(time = time + min(sir_data$time))

posterior

```

```{r}

posterior %>%
    gather(S:R, key = 'variable', value = 'value') %>%
    mutate(variable = factor(variable, levels = var_names)) %>%
    ggplot(aes(x = time, y = value, group = sims, colour = variable)) +
    geom_line(alpha = 0.01) +
    facet_wrap(~variable, scales = 'free_y')

```

```{r}

sir_sample <- filter(sir_data, time < as.Date("2020-04-01"))

# Prior simulation
sir_model <- stan(
    file = here("analysis/tmp_stochastic_model.stan"),
    data = list(
        N = 5000000,
        T = nrow(sir_data)
    ),
    iter = 2000,
    chains = 4,
    algorithm = 'Fixed_param'
)

sir_model

```
